### 2013-01-19 edupar-13/landslide.tex
shrink threadfork picture
Add SMP4 paragraph and clean up various stuff
landslide.tex: A couple initial text squeezes.

Landslide is an experimental tool for helping kernel authors find and debug concurrency errors.
It is based on the technique of systematic exploration~\cite{verisoft}, a way of exploring the state space of different possible thread interleavings in a concurrent system.
It follows in the footsteps of related tools such as
\shortversion{
CHESS~\cite{chess},
}{
CHESS~\cite{chess}, dBug~\cite{dbug-ssv}, and DeMeter~\cite{demeter},
}
but is the first tool we know of to apply systematic exploration in a kernel development environment.

Conceptually,
Landslide is designed as follows.
Students annotate their code so Landslide knows
which kernel thread is currently running.
After one kernel thread has run for some time,
Landslide triggers artificial clock interrupts
to force a new thread to interleave with the current one.
When a test program finishes execution according
to one pattern of thread switches,
Landslide rewinds the kernel's state
and resumes the test according to a different thread interleaving.
After each instruction,
Landslide applies several bug-detection predicates to
the kernel's state to detect
illegal heap accesses,
deadlock, infinite loops, and panics.
\shortversion{}{
In theory,
by forcing a thread switch after every non-scheduler instruction,
Landslide could apply its bug-detection predicates to every reachable
execution state.
Because this would require prohibitive time to complete,
in practice Landslide uses a variety of techniques to
thread-switch less often and to avoid repeating
bug-equivalent execution paths.
}
%Landslide is implemented as a module for Simics, the simulator our students use as an execution environment for their kernels. Simics can load Landslide while booting a student's kernel, and calls into Landslide once per simulated instruction and once per simulated memory access.
%Landslide uses this information to update its internal representation of the kernel's state, which it in turn uses to decide how to control the kernel's execution.

@@

Figure~\ref{fig:threadfork} shows code containing a timer-dependent bug common in many Pebbles implementations which we will use as a running example in the rest of this section.
This code is a simplified implementation of the \x{thread_fork} system call: it constructs data structures necessary for the new thread, then asks the scheduler to make the new thread runnable, then returns its assigned thread ID.
However, if the timer preempts execution at line~\ref{line:hax}, the child thread might run long enough to invoke \x{exit}, causing \x{child} to become a dangling pointer.

@@

\begin{figure}[t]
\includegraphics[width=0.48\textwidth]{threadfork/threadfork.pdf}
\caption{The state space of possible thread interleavings can be viewed as an {\em execution tree}.
Landslide uses timer interrupts to force different threads to run as it explores this tree.
If a kernel has concurrency errors, they will show up in some branches, but not others.
This tree shows the bug from Figure~\ref{fig:threadfork}, with operations from the parent thread shaded.
}
\label{fig:tree}
\end{figure}

@@

The strategy we describe above may cause Landslide to identify decision points in unrelated parts of the kernel, such as when accessing mutexes in unrelated and/or already-trusted system calls.
We provide interface options in \x{config.landslide} for the student to view currently identified decision points and to selectively eliminate them.
If a student were testing thread death and reaping, they might want decision points to appear in \x{wait} and \x{vanish} but not if unrelated virtual memory operations are also in progress.
They should write \x{within_function wait vanish} and \x{without_function} \x{destroy_address_space}.
The \x{within_function} directive requires that at least one of the specified functions shall be on the call stack when decision points are identified, and \x{without_function} requires the opposite.

>>>>

Landslide is an experimental tool for helping kernel authors find and debug concurrency errors.
It is based on the technique of systematic exploration~\cite{verisoft}, a way of exploring the state space of different thread interleavings in a concurrent system.
It follows in the footsteps of related tools such as
\shortversion{
CHESS~\cite{chess},
}{
CHESS~\cite{chess}, dBug~\cite{dbug-ssv}, and DeMeter~\cite{demeter},
}
but is the first tool we know of to apply systematic exploration in a kernel development environment.

Conceptually,
Landslide is designed as follows.
Students annotate their code so Landslide knows
which kernel thread is currently running.
After one kernel thread has run for some time,
Landslide triggers artificial clock interrupts
% to force a new thread to interleave with the current one.
to force the scheduler to run a different thread.
When a test program finishes execution according
to one pattern of thread switches,
Landslide rewinds the kernel's state
and resumes the test according to a different thread interleaving.
After each instruction,
Landslide applies several bug-detection predicates to
the kernel's state to detect
illegal heap accesses,
deadlock, infinite loops, and panics.
\shortversion{}{
In theory,
by forcing a thread switch after every non-scheduler instruction,
Landslide could apply its bug-detection predicates to every reachable
execution state.
Because this would require prohibitive time to complete,
in practice Landslide uses a variety of techniques to
thread-switch less often and to avoid repeating
bug-equivalent execution paths.
}
%Landslide is implemented as a module for Simics, the simulator our students use as an execution environment for their kernels. Simics can load Landslide while booting a student's kernel, and calls into Landslide once per simulated instruction and once per simulated memory access.
%Landslide uses this information to update its internal representation of the kernel's state, which it in turn uses to decide how to control the kernel's execution.

@@

Figure~\ref{fig:threadfork} shows code containing a timer-dependent bug common in many Pebbles implementations which we will use as a running example\shortversion{.}{in the rest of this section.}
This code is a simplified implementation of the \x{thread_fork} system call: it constructs data structures necessary for the new thread, then asks the scheduler to make the new thread runnable, then returns its assigned thread ID.
However, if the timer preempts execution at line~\ref{line:hax}, the child thread might run long enough to exit, causing \x{child} to become a dangling pointer.

@@

\begin{figure}[t]
	\centering
\includegraphics[width=0.42\textwidth]{threadfork/threadfork.pdf}
\caption{The state space of possible thread interleavings can be viewed as an {\em execution tree}.
Landslide uses timer interrupts to force different threads to run as it explores this tree.
If a kernel has concurrency errors, they will show up in some branches, but not others.
This tree shows the bug from Figure~\ref{fig:threadfork}, with operations from the parent thread shaded.
}
\label{fig:tree}
\end{figure}

@@

The strategy we describe above may cause Landslide to identify decision points in unrelated parts of the kernel, such as when accessing mutexes in unrelated and/or already-trusted system calls.
We provide interface options in \x{config.landslide} for the student to view currently identified decision points and to selectively eliminate them.
If a student were testing thread death and reaping, they might want decision points to appear in \x{wait} and \x{vanish} but not if unrelated virtual memory operations are also in progress.
Accordingly, they could write \x{within_function wait vanish} and \x{without_function} \x{destroy_address_space}.
The \x{within_function} directive requires that at least one of the specified functions shall be on the call stack when decision points are identified, and \x{without_function} requires the opposite.
